---
title: "Getting started with zhncommandR"
output:
  rmarkdown::html_vignette:
    toc: true
    css: themakr-vignette.css
vignette: >
  %\VignetteIndexEntry{Getting started with zhncommandR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7, fig.height = 4.5
)
library(zhncommandR)
```

`zhncommandR` packages the **ZHN Auditor** dashboard — a Shiny app for live
evaluation of haematological tumour-centre cohorts — together with the
backend helpers that drive it. The dashboard is launched with
`zhn_run_app()`. The rest of this vignette shows the same analyses run
programmatically against the bundled synthetic example.

## Example data

The package ships a 100 % synthetic example workbook with three sheets:
`Basisdaten`, `Komplexe Chemotherapie`, `Komplexe Diagnostik`.

```{r}
path <- zhn_example_path()
basename(path)
```

## Reading the cohort

```{r}
cohort <- zhn_read_cohort(path, verbose = FALSE)
cohort |>
  dplyr::select(dplyr::any_of(c(
    "name", "diagnose", "behandlungsjahr", "primaerfall",
    "psychoonkologie", "pfs", "os"
  ))) |>
  dplyr::glimpse()
```

`zhn_read_cohort()` calls `janitor::clean_names()`, drops empty `none*`
columns, and derives `behandlungsjahr` from the first matching date column.

## Quality indicators

The indicator table on the dashboard is just a `dplyr` summary of a fixed
indicator list:

```{r}
indicators <- c("tumorkonferenz", "psychoonkologie", "sozialdienst",
                "hiv_hepatitis")

cohort |>
  dplyr::select(dplyr::any_of(indicators)) |>
  tidyr::pivot_longer(
    dplyr::everything(),
    names_to = "Indikator", values_to = "Wert"
  ) |>
  dplyr::group_by(Indikator) |>
  dplyr::summarise(
    Positiv = sum(zhncommandR:::.as_yesno(Wert) %in% TRUE, na.rm = TRUE),
    Gesamt  = dplyr::n(),
    Anteil  = round(Positiv / Gesamt, 2),
    .groups = "drop"
  )
```

## OPS-8-544 therapy blocks

```{r}
therapy_raw <- zhn_read_therapy(path, verbose = FALSE)
blocks <- zhn_prepare_therapy_blocks(therapy_raw)
blocks |>
  dplyr::select(patient, therapieprotokoll, diagnose, datum) |>
  dplyr::slice_head(n = 6)
nrow(blocks)
```

## Kaplan-Meier (PFS)

```{r km, message = FALSE, warning = FALSE}
event_col <- if ("rezidiv_event" %in% names(cohort)) {
  cohort$rezidiv_event
} else if ("rezidiv" %in% names(cohort)) {
  cohort$rezidiv
} else {
  NULL
}
km <- data.frame(
  time  = suppressWarnings(as.numeric(cohort$pfs)),
  event = zhncommandR:::.as_event01(event_col, mode = "auto")
) |>
  dplyr::filter(!is.na(time), !is.na(event))

fit <- survival::survfit(survival::Surv(time, event) ~ 1, data = km)

# Tidy the survfit step function, then draw it with ggplot2 (start at (0, 1)).
km_curve <- dplyr::bind_rows(
  data.frame(time = 0, surv = 1),
  data.frame(time = fit$time, surv = fit$surv)
)
ggplot2::ggplot(km_curve, ggplot2::aes(x = time, y = surv)) +
  ggplot2::geom_step() +
  ggplot2::scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  ggplot2::labs(x = "Monate", y = "PFS", title = "Beispiel-PFS (synthetisch)") +
  ggplot2::theme_minimal(base_size = 12)
```

## Oncoprint table

```{r}
onco <- zhn_parse_oncoprint(cohort)
onco |>
  dplyr::select(patient_label, diagnose_label, alteration, alteration_class) |>
  dplyr::slice_head(n = 6)
```

## Launching the dashboard

```r
zhn_run_app()
```

## Use of LLM tools

Large language model tooling assisted with narrowly defined, non-authorial
tasks only: copyediting, prose smoothing, Markdown/LaTeX formatting,
scaffolding of boilerplate files (CI configs, build scripts), and code
refactoring. The tools were Chat AI (the LLM service of KISSKI, GWDG) and a
self-hosted Mistral Small (24B, Apache-2.0) run locally via Ollama and the
`ollamar` R package — local inference only, with no data sent to third
parties for the self-hosted model.

All scientific claims, methodological choices, analyses, interpretations, and
conclusions are the author's own. No LLM-generated text was incorporated
without review and revision, and every reference was verified against its
DOI, arXiv ID, or ISBN.
